---
title: "Week4-assignment"
author: "Miriam Kretschmer"
date: "21 3 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(ensembldb)
  library(epiwraps)
})
ah <- AnnotationHub()
```
```{r, eval = FALSE}
setwd("C:\\Users\\mkretschmer\\Documents\\Courses\\Bioinformatic_approaches\\Week4")
```


## Import peak-level data for Histone marks H3K4me3, H3K4me1, H3K27ac and H3K27me3 in mESC 
```{r}
peaks_p300 <- rtracklayer::import("C:\\Users\\mkretschmer\\Documents\\Courses\\Bioinformatic_approaches\\Week3\\peaks\\mouseChIP_peaks.narrowPeak", format = "narrowPeak")
peaks_H3K4me3 <- import("peaks/ENCFF055QNY.bed.gz", format = "narrowPeak")
peaks_H3K4me1 <- import("peaks/ENCFF284QUZ.bed.gz", format = "narrowPeak")
peaks_H3K27ac <- import("peaks/ENCFF587LEB.bed.gz", format = "narrowPeak")
peaks_H3K27me3 <- import("peaks/ENCFF055QNY.bed.gz", format = "narrowPeak")
```

How do they look like?
```{r}
plotSignalTracks(list(binned="C:\\Users\\mkretschmer\\Documents\\Courses\\Bioinformatic_approaches\\Week3\\tracks\\mouseChIP.bw", peaks=peaks_p300), region=peaks_p300[15], extend = 1000)
```
```{r}
head(peaks_H3K4me3)
head(peaks_H3K4me1)
head(peaks_H3K27ac)
head(peaks_H3K27me3)
```


## What proportion of p300 peaks overlap with the histone marks? 
```{r}
overlap_H3K4me3 <- overlapsAny(peaks_p300, peaks_H3K4me3)
table(overlap_H3K4me3)
overlap_H3K4me1 <- overlapsAny(peaks_p300, peaks_H3K4me1)
table(overlap_H3K4me1)
overlap_H3K27ac <- overlapsAny(peaks_p300, peaks_H3K27ac)
table(overlap_H3K27ac)
overlap_H3K27me3 <- overlapsAny(peaks_p300, peaks_H3K27me3)
table(overlap_H3K27me3)
```
None overlap.


# How long did the excercise take?
It was a bit tricky and took some google searching to figure the proper way to import the .bed.gz files. Besides that,  I was surprised to see that there should be no overlaps of any of the marks, coming from the literature I assumed there would be some. It took me about 1,5 h to finish this excercise.

```{r}
sessionInfo()
```